---
title: "NYPD Shooting incident data analysis"
date: '5/21/2021'
output:
  html_document: default
  pdf_document: default
---
****
# Introduction

### purpose

In this article, I will **clean**, **visualize** and **analyze** [NYPD Shooting incident data](https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic) to make it easier to see the trend of cumulative shooting incidents and murder cases per boroughs.  

### contents

1. Data Cleaning 
    i. Extract only the data needed for analysis
    ii. Date format conversion
    iii. Replace logical boolean data with integer type
    iv. Add the number of shooting occurrences and the cumulative number of it as columns.
    v. Add the cumulative number of murders as a column.
    
2. Visualization 
    a. Cumulative number of shooting events according to the flow of the date by borough
    b. Cumulative number of murder according to the flow of the date by borough  
    c. Percentage of murders in total number of shootings by borough
    
3. Analysis
    - Compare the total number of shooting cases and murder ones *by borough*. 
    - Calculate the percentage of murders in shootings *by borough*
      + `BROOKLYN`
      + `BRONX`
      + `QUEENS`
      + `MANHATTAN`
      + `STATEN ISLAND`
      
4. Model 
    - Compare the actual trend of shooting incidents with the predictive linear model
    
    
5. Conclusion and Bias Identification   
    - Conclude the project report 
    - Identify personal bias and mitigation method
    
    
<br /> 
<br /> 
    
    
    
    

### 1.Data Cleaning

Data used in this project is every shooting incident data list occurred in NYC from Jan. 1, 2006 to Dec. 31, 2020. 



```{r pressure, echo=FALSE, fig.height=5, fig.width=5}
# plot(pressure)
```



**Data Importing**
<br />
Here you can see the original dataset below. 

```{r}
library(knitr)
library(tidyr)
library(tidyverse)
library(lubridate)

nypd <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
knitr::kable(head(nypd))
```

**Data cleaning**
<br />
Date format was converted for easier reading and I replaced logical boolean data, 'STATISTICAL_MURDER_FLAG' with integer type like '0' for FALSE and '1' for TRUE. And I added the number of shooting occurrences and the cumulative number of it as columns with the names of 'shooting' and 'cumshooting'. Also a column, 'cummurder', was added for the cumulative number of murders. 
```{r}
library(dplyr)
library(ggplot2)


# select only needed data
nypd_test <- drop_na(nypd) %>%
  select(-c(INCIDENT_KEY, LOCATION_DESC, X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat, JURISDICTION_CODE, OCCUR_TIME))

# change the date type 
nypd_test <- nypd_test %>%
mutate(OCCUR_DATE = mdy(OCCUR_DATE))

#change logical boolean into int.
nypd_test$STATISTICAL_MURDER_FLAG [nypd_test$STATISTICAL_MURDER_FLAG == "TRUE"] <- 1
nypd_test$STATISTICAL_MURDER_FLAG [nypd_test$STATISTICAL_MURDER_FLAG == "FALSE"] <- 0



nypd_murder_boro <- nypd_test %>%
     group_by(BORO) %>%
#     summarize(STATISTICAL_MURDER_FLAG = sum(STATISTICAL_MURDER_FLAG)) %>%
     select(BORO, OCCUR_DATE, STATISTICAL_MURDER_FLAG) %>%
     ungroup()

nypd_murder_boro_1 <- nypd_murder_boro %>%
     group_by(BORO, OCCUR_DATE) %>%
     summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, STATISTICAL_MURDER_FLAG) %>%
     ungroup()

# add new columns
nypd_murder_boro_1$cummurder <- ave(nypd_murder_boro_1$STATISTICAL_MURDER_FLAG, nypd_murder_boro_1$BORO, FUN = cumsum)

nypd_murder_boro_1['shooting'] = 1

nypd_murder_boro_1$cumshooting <- ave(nypd_murder_boro_1$shooting, nypd_murder_boro_1$BORO, FUN = cumsum)

nypd_murder_boro_1$murderpercent <- with(nypd_murder_boro_1, cummurder/cumshooting *100)
# show the data that will be used for analysis
knitr::kable(head(nypd_murder_boro_1))
```


<br />
<br />

### 2. Visualization 
Here comes three graphs. The first two shows the cumulative number of murder and of shooting events according to the flow of the date by borough. And the rest shows the percentage of murders in total number of shootings by borough. 
```{r}
#Visualization

nypd_murder_boro_1 %>%
  ggplot(aes(x = OCCUR_DATE, y=cummurder, group=BORO, color=BORO))+
  geom_line()


nypd_murder_boro_1 %>%
  ggplot(aes(x = OCCUR_DATE, y=cumshooting, group=BORO, color=BORO))+
  geom_line()


nypd_murder_boro_1 %>%
  ggplot(aes(x = OCCUR_DATE, y=murderpercent, group=BORO, color=BORO))+
  geom_line()

```


<br />
<br />

### 3. Analysis
Comparing the total number of shooting cases and murder ones by borough I calculated the percentage of murders in shootings and found that in STATEN ISLAND the total number of shootings is the lowest, but the proportion of deaths from shootings is the highest. 


```{r}
# analysis

aggregate(nypd_murder_boro_1$STATISTICAL_MURDER_FLAG, by=list(BORO = nypd_murder_boro_1$BORO), FUN=sum)
aggregate(nypd_murder_boro_1$shooting, by=list(BORO = nypd_murder_boro_1$BORO), FUN=sum)

city <- "BRONX"
nypd_murder_boro_BRONX <- nypd_murder_boro_1 %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_BRONX))

city <- "BROOKLYN"
nypd_murder_boro_BROOKLYN <- nypd_murder_boro_1 %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_BROOKLYN))


city <- "STATEN ISLAND"
nypd_murder_boro_STATENISLAND <- nypd_murder_boro_1 %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_STATENISLAND))

city <- "MANHATTAN"
nypd_murder_boro_MANHATTAN <- nypd_murder_boro_1 %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_MANHATTAN))

city <- "QUEENS"
nypd_murder_boro_QUEENS <- nypd_murder_boro_1 %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_QUEENS))




```


<br />
<br />

### 4. Model 
I made a linear model to predict the shooting incident of boroughs and compared it to the actual trend of shooting incidents. In graph below, the blue line represents the actual trend of shooting incidents and the red does the predictive model prediction. 
```{r}


# Modeling Data

mod <- lm(cumshooting ~ cummurder, data = nypd_murder_boro_1)
summary(mod)


nypd_murder_boro_1 %>% slice_min(cumshooting)
nypd_murder_boro_1 %>% slice_max(cumshooting)

x_grid <- seq(0, 3000)
new_df <- tibble(cumshooting = x_grid)

nypd_pred <- nypd_murder_boro_1 %>% mutate(pred = predict(mod))

# nypd_pred

nypd_pred %>% ggplot() +
  geom_point(aes(x = OCCUR_DATE, y=cumshooting), color= "blue")+
  geom_point(aes(x = OCCUR_DATE, y = pred), color = "red")

```


<br />
<br />

### 5. Conclusion and Bias Identification


The analysis shows that BROOKLYN is the place where the most shootings occurred in the data. However, STATEN ISLAND has the highest rate of deaths from shooting. If social policy is established based on this data, I think that prevention education for shootings should be approached in a different way in BROOKLYN and STATEN ISLAND.
<br />
Born in a country in which possession of firearms itself is illegal, I have great fear of shooting itself and distrust of the society in which it is carried out.
So, rather than seeing and understanding the data of citizens who can legally own firearms as an important part of society and thinking about countermeasures, the fact that there are many gunshots is just bad.
However, while doing science, it has changed that I try to accept the phenomenon itself and recognize the value of research.
Not making judgments about certain facts but trying to accept and understand the phenomenon in a neutral way is a way for me to step into a bigger world, and for our society to make the rules fairer. 

